---
title: "R Notebook"
output: html_notebook
---

 

```{r}
library(stringr)
library(tidyverse)
library(lubridate)
```

Load data and correct column name. 
```{r}
arb.figures <- read.csv("./Data/SalaryArbitration_MLBTR.csv", stringsAsFactors = F)
colnames(arb.figures)[1] <- 'Player'
arb.figures <- na.omit(arb.figures)
```

Split player names into first and last names
```{r}
arb.figures$nameFirst <-str_split_fixed(arb.figures$Player, " ", 2)[,1]
arb.figures$nameLast <- str_split_fixed(arb.figures$Player, " ", 2)[,2]
# fix names for merging with Lahman database
arb.figures[arb.figures$nameFirst == "A.J.", 'nameFirst'] <- "A. J."
arb.figures[arb.figures$nameFirst == "C.J.", 'nameFirst'] <- "C. J."
arb.figures[arb.figures$nameFirst == "J.A.", 'nameFirst'] <- "J. A."
arb.figures[arb.figures$nameFirst == "J.C.", 'nameFirst'] <- "J. C."
arb.figures[arb.figures$nameFirst == "J.D.", 'nameFirst'] <- "J. D."
arb.figures[arb.figures$nameFirst == "J.J.", 'nameFirst'] <- "J. J."
arb.figures[arb.figures$nameFirst == "J.P.", 'nameFirst'] <- "J. P."
arb.figures[arb.figures$nameFirst == "J.T.", 'nameFirst'] <- "J. T."
arb.figures[arb.figures$nameFirst == "T.J.", 'nameFirst'] <- "T. J."
arb.figures[arb.figures$nameFirst == "Yolmer", 'nameFirst'] <- "Carlos"
arb.figures[arb.figures$Player == 'Lance McCullers Jr.', c('nameFirst')] <- 'Lance'
arb.figures[arb.figures$Player == 'Lance McCullers Jr.', c('nameLast')] <- 'McCullers'
arb.figures[arb.figures$Player == 'Jonathan Gray', c('nameFirst')] <- 'Jon'
arb.figures[arb.figures$Player == 'Matt Joyce', c('nameFirst')] <- 'Matthew'
arb.figures[arb.figures$nameLast == "D'Arnaud", 'nameLast'] <- "d'Arnaud"
arb.figures[arb.figures$nameLast == "Milone", 'nameFirst'] <- "Tommy"
arb.figures[arb.figures$nameFirst == "Zack" & arb.figures$nameLast == 'Wheeler', 'nameFirst'] <- "Zach"
arb.figures[arb.figures$nameLast == "de Aza", 'nameLast'] <- "De Aza"
arb.figures[arb.figures$nameLast == "Edwards Jr.", 'nameLast'] <- "Edwards"
arb.figures[arb.figures$nameLast == "DeShields Jr.", 'nameLast'] <- "DeShields"
arb.figures[arb.figures$nameLast == "Bradley Jr.", 'nameLast'] <- "Bradley"
arb.figures[arb.figures$nameLast == "Souza Jr.", 'nameLast'] <- "Souza"
arb.figures[arb.figures$nameFirst == "Felipe" & arb.figures$nameLast == 'Vazquez', 'nameLast'] <- "Rivero"
arb.figures[arb.figures$nameFirst == "John" & arb.figures$nameLast == 'Ryan Murphy', c('nameFirst', 'nameLast')] <- c('J. R.', 'Murphy')
arb.figures[arb.figures$nameLast == "Upton, Jr.", 'nameLast'] <- "Upton"
arb.figures[arb.figures$Player == 'Michael Dunn', c('nameFirst')] <- 'Mike'
arb.figures[arb.figures$Player == 'Michael Foltynewicz', c('nameFirst')] <- 'Mike'
arb.figures[arb.figures$Player == 'Michael Fiers', c('nameFirst')] <- 'Mike'
arb.figures[arb.figures$Player == 'Nate Eovaldi', c('nameFirst')] <- 'Nathan'
arb.figures[arb.figures$Player == 'Nicholas Castellanos', c('nameFirst')] <- 'Nick'
arb.figures[arb.figures$Player == 'Samuel Freeman', c('nameFirst')] <- 'Sam'
arb.figures[arb.figures$Player == 'Norichika Aoki', c('nameFirst')] <- 'Nori'
arb.figures[arb.figures$Player == 'Juan Carlos Oviedo', c('nameFirst', 'nameLast')] <- c('Juan Carlos', 'Oviedo')

```

Merge with the lahman master file for playerIDs.
```{r}
# Load lahman database
lahman.master <- read.csv('./Data/Lahman/People.csv') 
# remove players who debuted before 200 to minimize duplicates
lahman.master <- lahman.master[!(is.na(lahman.master$debut)),]
lahman.master$debut <- as.Date(lahman.master$debut, "%Y-%m-%d")
lahman.master$birthDate <- as.Date(lahman.master$birthDate, "%Y-%m-%d")

# Filter old players for the merging with first and last names
lahman.master <- lahman.master %>% 
  filter(year(debut) > 1995)

# Merge arbitration and lahman data
arb.master <- left_join(arb.figures, lahman.master, by = c('nameFirst', 'nameLast'))

```

Now remove the duplicates 
```{r}
# Find IDs of players who got added because they have the same first and last name as a player in the arbitration
# data
wrong.IDs <- data.frame(bbrefID = c('carpeda02', 'cartech01', 'duffyma02', 'cruzne01',
                                    'fernajo01', 'fieldjo02', 'garcilu01', 'garcilu02',
                                    'gonzami04', 'gonzami05', 'hatchch01', 'martica03', 
                                    'ramirer01', 'ramirra03', 'reynoma03', 'rodrifr03',
                                    'taylomi01', 'fernajo03', 'fernajo04', 'guerrja02'))

# Remove the erroneously added players
arb.master <- arb.master[!(arb.master$bbrefID %in% wrong.IDs$bbrefID),] 

```


Get previous year salary
```{r}
# LOad Lahman salary data and subtract previous year to get current yet
lahman.salaries <- read.csv('./Data/Lahman/Salaries.csv') 
arb.master$prev.year <- arb.master$Year - 1

# Get previous salary
prev.salary <- left_join(arb.master, 
                        lahman.salaries[,c('yearID', 'playerID', 'salary')],
                        by = c('bbrefID' = 'playerID', 'prev.year' = 'yearID'))
arb.master$prev.salary <- prev.salary$salary

#select limited columns
arb.master2 <- arb.master[, c(1:14, 18, 28:32, 36:37)]
sum(is.na(arb.master2$prev.salary))
```
That are a lot of missing values


Add baseball reference scraped salaries to existing data frame if values are missing for the current value
```{r}
# read in BR data
BR.salaries <- read.csv("./Data/PlayerSalaryBR.csv", stringsAsFactors = F)

# get playerID for BR.salaries
#BR.salaries <- left_join(BR.salaries, lahman.master[,c('bbrefID', 'birthDate', 'nameFirst', 'nameLast')], 
#                         by = c('bbrefID' = 'playerID'))

# Add BR salaries to dataframe
arb.master.salaries <- arb.master2 %>%
  left_join(BR.salaries[,c('Salary', 'year', 'bbrefID')],
            by = c('playerID' = 'bbrefID', c('prev.year' = 'year')))

# Set NAs to 0 for comparison 
arb.master.salaries[is.na(arb.master.salaries$Salary) | arb.master.salaries$Salary == "",'Salary'] <- 0
arb.master.salaries[is.na(arb.master.salaries$prev.salary) | arb.master.salaries$prev.salary == "",'prev.salary'] <- 0

# Remove symbols
arb.master.salaries$Salary <- gsub('\\$|,', "", arb.master.salaries$Salary)

# Add BR salaries if exisiting previous salary is missing
arb.master.salaries[arb.master.salaries$prev.salary == "0", 'prev.salary'] <- 
  arb.master.salaries[arb.master.salaries$prev.salary == "0", 'Salary']

# Remove BR salaries from df
arb.master.final <- arb.master.salaries[,-25]

# Find number of missing salaries
sum(arb.master.final$prev.salary == 0)

```
Still have missing salaries


Combine and clean up salary info. 
```{r}
# Remove MM and $, change to numeric and convert to millions
arb.master.final$Player.Amt.<- as.numeric(gsub('\\$|M', "", arb.master.final$Player.Amt.))*1000000
arb.master.final$Team.Amt.<- as.numeric(gsub('\\$|M', "", arb.master.final$Team.Amt.))*1000000
arb.master.final$Midpoint<- as.numeric(gsub('\\$|M', "", arb.master.final$Midpoint))*1000000
arb.master.final$Prev.Salary<- as.numeric(gsub('\\$|M', "", arb.master.final$Prev.Salary))*1000000

# Further check existing salaries 
arb.master.final[is.na(arb.master.final$Player.Amt.), 'Player.Amt.'] <- 0
arb.master.final[is.na(arb.master.final$Team.Amt.), 'Team.Amt.'] <- 0
arb.master.final[is.na(arb.master.final$Midpoint), 'Midpoint'] <- 0
arb.master.final[is.na(arb.master.final$Prev.Salary), 'Prev.Salary'] <- 0
arb.master.final[arb.master.final$Prev.Salary == 0, "Prev.Salary"] <-
  arb.master.final[arb.master.final$Prev.Salary == 0, 'prev.salary']
arb.master.final <- arb.master.final[,-24]

sum(arb.master.final$Prev.Salary == 0)
```
Less missing values

# Add unknown valeus
```{r}
# See characteristics the players with missing salaries
arb.master.final[arb.master.final$Prev.Salary == 0, 1:5]
# Many have Super 2 designations, let's get minimum salaries by year 

# Get super minimum salaries
min.salaries <- data.frame('Year' = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019),
                           'Salary.Min' = 1000*c(414, 480, 490, 500, 507.5, 507.5, 535, 545, 555))

# Assign minimum salaries to super 2s
arb.master.final <- inner_join(arb.master.final, min.salaries, by = c('prev.year' = 'Year'))
arb.master.final[arb.master.final$Prev.Salary == 0 & arb.master2$Super2 == "Yes", 'Prev.Salary'] <-
  arb.master.final[arb.master.final$Prev.Salary == 0 & arb.master2$Super2 == "Yes", 'Salary.Min'] 

# Assign minimum salaries to players with less than 3.16 years of service
arb.master.final[arb.master.final$Prev.Salary == 0 & arb.master.final$Service < 3.2, 'Prev.Salary'] <- 
  arb.master.final[arb.master.final$Prev.Salary == 0 & arb.master.final$Service < 3.2, 'Salary.Min'] 

arb.master.final[arb.master.final$Prev.Salary == 0, c('Player', 'Year', 'Team', 'Service', 'Prev.Salary')]
```
Much better

Now these will have to be done manually
```{r}
# Get previous salaries for players with missing values
arb.master.final[arb.master.final$Prev.Salary == 0, 'Prev.Salary'] <- c(507500, 507500, 507500,
                                                                        507500, 2000000, 500000,
                                                                        490000, 545000, 9715847,
                                                                        414000, 750000, 557500,
                                                                        1625000, 414000, 480000, 
                                                                        545000, 507500, 625000,
                                                                        507500, 490000, 414000,
                                                                        1475000, 500000, 500000,
                                                                        2050000, 2500000)

arb.master.final$Prev.Salary<- as.numeric(gsub('\\$|M', "", arb.master.final$Prev.Salary))*1000000

```

Remove characters from settled amount with a dollar value
```{r}
# Find the values with Extension or Unknown
masker <- !(arb.master.final$Settled.Amt. == 'Extension' | arb.master.final$Settled.Amt. == 'Unknown' )
# Remove the values
arb.master.final[masker ,'Settled.Amt.'] <-
  as.numeric(gsub('\\$|M', "",arb.master.final[masker, 'Settled.Amt.']))*1000000 
```


```{r}
write.csv(arb.master.final, file = "./Data/arbitration_contracts.csv")

```
